Tracing the Meta-Level: PyPy's JIT Compiler

### 2009-03-31 talk/icooolps2009/paper.tex
Accepting one of Anto's changes, rejecting the rewrite: The problem with the rewrite was that it jumped to the conclusion of where the text gets to several paragraphs later. Instead, tried to rewrite it slightly to make it more clear.

\anto{I find this para a bit confusing, but I don't really know how to improve
  it. I try a rewriting, feel free to revert}

\sout{A tracing JIT compiler finds the hot loops of the program it is compiling. In
our case, this program is the language interpreter. The hot loop of the language
interpreter is the bytecode dispatch loop. Usually it is also the only hot loop.
Tracing one iteration of this loop means that the execution of one opcode was
seen. This means that the resulting machine code will correspond to a loop, that
assumes that this particular opcode will be executed many times in a row,
which is clearly very unlikely.}

\anto{Similarly, we need to distinguish two kind of loops: \emph{interpreter
    loops} are loops \textbf{inside} the language interpreter (e.g., the
  bytecode dispatch loop). \emph{User loops} are loops in the user program.}

\anto{To improve the performances of the user programs, we need to find and
  compile traces that correspond to user loops.  The strategy described in
  section \ref{sect:tracing} does not work well, as it only detects
  interpreter loops, and in particular the bytecode dispatch one.}

@@

\anto{ Because of the guard on \texttt{opcode0}, the code compiled from this
  trace will be useful only when executing a \texttt{DECR\_A}.  For all the
  other operations the guard will fail, making the whole system probably
  slower than without JIT.}

>>>>

Similarly, we need to distinguish loops at two different levels:
\emph{interpreter loops} are loops \emph{inside} the language interpreter. On
the other hand, \emph{user loops} are loops in the user program.

A tracing JIT compiler finds the hot loops of the program it is compiling. In
our case, this program is the language interpreter. The hot loop of the language
interpreter is its bytecode dispatch loop. Usually that is is also the only hot
loop of the language interpreter.  Tracing one iteration of this loop means that
the recorded trace corresponds to execution of one opcode. This means that the
assumption that the tracing JIT makes -- that several iterations of a hot loop
take the same or similar code paths -- is just wrong in this case. It is very
unlikely that the same particular opcode is executed many times in a row.

@@

Because of the guard on \texttt{opcode0}, the code compiled from this trace will
be useful only when executing a long series of \texttt{DECR\_A} opcodes. For all
the other operations the guard will fail, which will mean that performance is
probably not improved at all.
